---
title: "Example model run"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{run-model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(sf)
library(mapview)
mapviewOptions(
  fgb = FALSE,
  vector.palette = colorRampPalette(
    c("#000004FF", "#420A68FF", "#932667FF", "#DD513AFF", "#FCA50AFF", "#FCFFA4FF")
  )
)
```

Tool for placing bicycle sharing stations by iteratively minimizing a global energy of a model that reflects the station-based bicycle sharing system. The station locations are randomly initialized in the street network and iteratively optimized based on the reachable population in combination with walking and driving times.

The model in the package example optimizes the positions of stations in an assumed bicycle sharing service for the municipality Bülach in Zurich, Switzerland.

## Setup
```{r }
library(drtplanr)
set.seed(123)
```

## Example data sets
Load the package example data sets:

* aoi: Area of Interest - boundary of the municipality Bülach in Zurich, Switzerland.
* pop: Centroids of population and structural business hectare grid statistis (BfS).
* poi: Points od Interest - manually selected points to place a (fixed) station.

```{r eval=TRUE, out.width='100%'}
aoi <- 
  sf::st_read(system.file("example.gpkg", package = "drtplanr"), layer = "aoi")

pop <- 
  sf::st_read(system.file("example.gpkg", package = "drtplanr"), layer = "pop")

poi <-
  sf::st_read(system.file("example.gpkg", package = "drtplanr"), layer = "poi")

#m <-
#  mapview(aoi, alpha.region = 0, layer.name = "AOI", homebutton = FALSE) +
#  mapview(pop, zcol = "n", alpha = 0, layer.name = "Population", homebutton = FALSE)
#m
```

## Initialize drtm model
Create a new model.
```{r }
m <- drt_drtm(
  model_name = "example",
  aoi = aoi, poi = poi, pop = pop,
  n_sta = 20, m_seg = 100
)
m
```

## Minimze the energy of the model
Iterate the model 100 times, where every iteration consists of:

1. Relocate a station randomly on the street segments.
2. Calculate the new global energy of the model using the routing graphs.
3. If the energy is lower than previuos iteration: Keep the new location of the virtual station; otherwise: Reset to the previous location.

```{r include=FALSE, eval=TRUE}
m1 <- drt_iterate(m, 10)
```

```{r eval=FALSE}
m1 <- drt_iterate(m, 10)
```

Print model summary:
```{r out.width='100%'}
m1
```

## Plotting
### Energy curve
```{r out.width='100%'}
drt_plot(m1)
```

### Station map
```{r eval=FALSE, out.width='100%'}
#drt_map(m1)
```

## Iterate again
The model state is saved in the model. This allows making copies and store different stages of the optimization.
```{r include=FALSE, out.width='100%'}
m2 <- drt_import(
  system.file("example_i1000.RData", package = "drtplanr")
)
```

```{r eval=FALSE, out.width='100%'}
m2 <- drt_iterate(m1, 990)
```

Print model summary:
```{r out.width='100%'}
m2
```

## Visualize results
```{r out.width='100%'}
drt_plot(m2)
```

```{r eval=FALSE, out.width='100%'}
#drt_map(m2)
```

## Export and import
The drtplanr has functions to import and export the drtm objects. Furthermore graphics of the energy plot and the station map can be exported as images (.png).
```{r eval=FALSE, out.width='100%'}
# Export model
drt_export(m2, path = getwd())

# Import model
m2 <- drt_import("example_i1000.RData")

# Save graphics
drt_save_graphics(m2, path = getwd())
```

## References
* [hereR](https://github.com/munterfinger/hereR): R interface to the HERE REST APIs 
* [BfS](https://www.bfs.admin.ch/): Population data for Switzerland
* [OSM](https://www.openstreetmap.org/): Street network data for routing purposes.

